---
title: "R Notebook"
output: html_notebook
---

NAME: Sravani Mahankali
FILE: Examination

---------------------------------------------------------------------------------------------

```{r}
## installing the necessary packages

install.packages("ggplot2")
install.packages("faraway")
install.packages("generalhoslem")
install.packages("tidyverse")
install.packages("brant")
install.packages("MASS")
```

```{r}
## loading the packages

packages <- c('survival','dplyr','ggplot2','GGally','faraway', 'generalhoslem', 'tidyverse','brant','MASS')
purrr::walk(packages,library,character.only=T)
```

```{r}
## loading the data and checking the summary

data(pbc)
?pbc
summary(pbc)
```

```{r}
## omiting id variable

my.pbc <- na.omit(pbc[, -1])
```

---------------------------------------------------------------------------------------------
1) The starter file has code for converting a couple of variables into factors, from their original type. What is the purpose behind these conversions? Base you answer in the contextual information related to the dataset. This information can be pulled up by using the help function on pbc.


```{r}
## converting variables to factors as they have categorical nature

my.pbc$status <- as.factor(my.pbc$status) #status should be a factor
my.pbc$ascites <- as.factor(my.pbc$ascites) #ascites should be factor
my.pbc$stage <- as.factor(my.pbc$stage)#stage is ordinal/rank
my.pbc$trt <- as.factor(my.pbc$trt)
my.pbc$hepato <- as.factor(my.pbc$hepato)
my.pbc$spiders <- as.factor(my.pbc$spiders)
my.pbc$edema <- as.factor(my.pbc$edema)
```

```{r}
## checking the summary

summary(my.pbc)
```

```{r}
## checking the class of all variables 

sapply(my.pbc, class)
```

```{r}
## checking the class of only stage

summary(my.pbc$stage)
```

---------------------------------------------------------------------------------------------
2) Execute the code associated with excluding trt, time, and status from the dataframe to produce a reduced version of the dataset (represented by the dataframe object labeled my.pbc in the R code file accompanying these instructions).


```{r}
## ignoring trt, time, and status, creating a reduced dataset

my.reduced.pbc <- my.pbc[, -c(1:3)]
summary(my.reduced.pbc)
```


---------------------------------------------------------------------------------------------

```{r}
## checking the pair plot to identify variables leading to data redundancy

ggpairs(my.reduced.pbc, upper = list(continuous = wrap("cor", size = 1.0)))+
  theme_minimal(base_size = 4)
```


```{r}
## ordinal regression full model

ordreg1 <-  MASS::polr(stage ~ ., my.reduced.pbc, Hess=TRUE)
summary(ordreg1)

```

```{r}
## choosing better ordinal regression model by removing un-necessary variables to reduce data redundancy

better.model <- step(ordreg1)
```

```{r}
## removing unnecessary variables

My.Reduced.Pbc <- my.reduced.pbc[,-c(2,6,7,9,11,12,13,14, 15)]

summary(My.Reduced.Pbc)
```

---------------------------------------------------------------------------------------------
2) Then, fit a model to predict stage using the remaining variables as predictors in the reduced version of the dataset (represented by the dataframe object with the label my.reduced.pbc). Explain the purpose behind excluding the above listed three variables as predictors. 

and

3) Explain the rationale behind the specific modeling approach you have used, on the basis of the measurement type of the outcome variable.

```{r}
## model after removing the variables to reduce data redundancy

ordreg <-  polr(stage ~ ., My.Reduced.Pbc, Hess = TRUE)
summary(ordreg)
```

```{r}
## calculating coeffitients summary

ctable <- coef(summary(ordreg))

## calculating p values

p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2

## combining p values and ctable

ctable <- cbind(ctable, "p value" = p)

ctable
```

```{r}
## confidence intervals

ci <- confint.default(ordreg)

## odds ratio

exp(cbind(OR = coef(ordreg), ci))
```

---------------------------------------------------------------------------------------------
4) State the key assumptions underlying your choice of the modeling technique. Then, perform the necessary diagnostic tests to determine the extent to which these assumptions were met.

```{r}
## multi-collinearity assumption

car::vif(ordreg)
```

```{r}
## proportional odds

brant::brant(ordreg)
```

```{r}
## hosmer-lemeshow test

generalhoslem::lipsitz.test(ordreg)
```

---------------------------------------------------------------------------------------------
5)Next, go back to using my.pbc and add a new variable to represent two states of status: the
original status values of 0 and 1 are mapped to a 0 and the original value of 2 onto a 1. In the R code, this new variable is labeled newstatus. What purpose would be served by combining the original 0 (censored) and 1 (transplant) into a single category, and the original 2 (dead) into another single category? (Hint: consider what you would need for status, if your goal is to model survival/non-survival). What is lost when combining the original 0 and 1 categories into a single new category? How will this affect the interpretation of the results.


```{r}
## creating a new variable, combining censored and transplant into one factor, and dead as another

my.pbc$newstatus <- as.factor(ifelse((my.pbc$status == 0),0,
                                ifelse((my.pbc$status == 1),0,
                                      1)))
 
summary(my.pbc$newstatus)
```

```{r}
## removing the original status variable from the dataset

my.pbc <- my.pbc[, -2]

summary(my.pbc)
```

---------------------------------------------------------------------------------------------

```{r}
## building the ggpairs to check the data redundancy issue due to high collinearity

ggpairs(my.pbc, upper = list(continuous = wrap("cor", size = 1.0)))+
  theme_minimal(base_size = 4)
```


```{r}
## building the CoxPH model with all the predictors

coxph.model1 <- coxph(Surv(time, as.numeric(newstatus)) ~  trt + age + sex + ascites + hepato                          + spiders + edema + bili + chol + albumin + copper + alk.phos + ast                          + trig + platelet+ protime + stage,
                          data=my.pbc)

summary(coxph.model1)

```

```{r}
## choosing better model to reduce data redundancy by removing unnecessary variables

best.model <- step(coxph.model1)

```

---------------------------------------------------------------------------------------------
6) Next, fit a CoxPH model on all of the predictors. Would it make sense to retain status in the set of predictors? Why/why not? Identify the variables that are indicated as being significant predictors of the outcome. What is the outcome that is being modeled?


```{r}
## building the better model

coxph.best <- coxph(Surv(time, as.numeric(newstatus)) ~  age  + edema + bili + albumin + copper+ ast 
                    + protime + stage,
                    data=my.pbc)
summary(coxph.best)

```

```{r}
## plotting the model

plot(survfit(coxph.best), ylim = c(0.2, 1.0), xlab="Days", ylab="Proportion not dead",
     main="Main survival function's plot" )
```

---------------------------------------------------------------------------------------------
7) Identify the key assumption(s) associated with the CoxPH modeling approach, and explain
whether it/they are satisfied, using appropriate evidence. Summarize your results in a table with the following structure.


```{r}
## checking proportional hazards assumption

cox.zph(coxph.best)
```


```{r}
## plotting proportional hazards of best model

par(mfrow=c(2, 3))
plot(cox.zph(coxph.best))

```

```{r}
### checking the influential observations

dfbeta <- residuals(coxph.best, type="dfbeta")

## plotting the influential observations

par(mfrow=c(2, 3))
for (j in 1:11) {
  plot(dfbeta[, j], ylab=names(coef(coxph.best))[j])
  abline(h=0, lty=2)
}
```

```{r}
## plotting the residuals and predictors linear plot to check linearity


plot(predict(coxph.best), residuals(coxph.best, type = "deviance"))
abline(h=0)
lines(smooth.spline(predict(coxph.best), residuals(coxph.best, type = "deviance")), col = "blue")

```

---------------------------------------------------------------------------------------------
8) Next, fit full-parametric models, where the survival function of status is modeled using
exponential and Weibull distributions to model the outcome (survival function of newstatus,
whereas the CoxPH approach models the hazard function of newstatus). Which of these is a
better model? Explain using appropriate evidence.


```{r}
## fitting weibull distribution

mod.weibull <- survreg(Surv(time, as.numeric(newstatus)) ~ age  + edema + bili + albumin + copper+ ast 
                        + protime + stage, data=my.pbc)
summary(mod.weibull)
```

```{r}
## fitting the exponential model

mod.exponential <- survreg(Surv(time, as.numeric(newstatus)) ~ age  + edema + bili + albumin + copper+ ast 
                           + protime + stage, data=my.pbc, dist="exponential")
summary(mod.exponential)
```

---------------------------------------------------------------------------------------------
9) Taking into account the results from the three survival-related models, identify the significant predictors of the outcome. Among these, which variable is the most important one? Explain on the basis of appropriate evidence.


```{r}
## choosing the better model among parametric models

AIC(mod.exponential, mod.weibull)
```

```{r}
## choosing the better model among all the three survival models

AIC(mod.exponential, mod.weibull, coxph.best)
```

---------------------------------------------------------------------------------------------



